---
title: "DSC 200 Lab Project -Term 2232"
date: "`r Sys.Date()`"
output: github_document
---

**Student Name:<insert your name here>**wajdan

**Student ID:<insert ID here>**2221003984


**Deadline:** 23:59 on Sunday, 19 May 2024

**Total Points:** 20


## Loading Packages

```{r load-packages, message = FALSE}

library(tidyverse)
library(openintro)
library(ggrepel)
```


## Tasks

`1. (2 points)
```{r}num_pets <- nrow(seattlepets)
num_pets

```
Write your narrative We use the nrow() function to count the number of rows in the seattlepets dataset, which gives us the number of pets included in the dataset. 




`2. (2 points)

```{r}num_variables <- ncol(seattlepets)
num_variables

```
Write your narrative below We use the ncol() function to count the number of columns in the seattlepets dataset, which gives us the number of variables for each pet.



`3. (2 points)
```{r}species_count <- seattlepets %>%
  count(species) %>%
  arrange(desc(n))
species_count

```
Write your narrative here We use the count() function to count the number of pets for each species in the seattlepets dataset. Then we arrange the results in descending order using the arrange(desc(n)) function.


`4. (2 points)
```{r}common_pet_names <- seattlepets %>%
  count(animal_name) %>%
  arrange(desc(n)) %>%
  top_n(10)
common_pet_names

```
Write your narrative here We use the count() function to count the frequency of each pet name in the seattlepets dataset. Then we arrange the results in descending order using the arrange(desc(n)) function and select the top ten most common names using the top_n(10) function.


`5. (2 points)
```{r}pig_records <- seattlepets %>%
  filter(species == "Pig") %>%
  arrange(animal_name)
pig_records

```
Write your narrative her eWe use the filter() function to filter the records where the species is "Pig". Then we arrange the results by pet names using the arrange(animal_name) function




`6. (2 points)
```{r}goat_records <- seattlepets %>%
  filter(species == "Goat") %>%
  select(animal_name, primary_breed) %>%
  arrange(animal_name)
goat_records

```
Write your narrative here We use the filter() function to select the records where the species is "Goat". Then we use the select() function to choose the columns animal_name and primary_breed. Finally, we sort the results by pet names using the arrange(animal_name) function





`7. (2 points)
```{r}merged_columns <- seattlepets %>%
  mutate(pet = paste(animal_name, species, sep = "; ")) %>%
  select(license_number, pet) %>%
  arrange(pet)
merged_columns

```
Write your narrative here We use the mutate() function to add a new column pet which concatenates the values from animal_name and species using the paste() function with a separator "; ". Then we use the select() function to choose the columns license_number and pet. Finally, we sort the results by the pet column using the arrange(pet) function.



`8. (2 points)
```{r}species_plot <- seattlepets %>%
  ggplot(aes(x = species)) +
  geom_bar() +
  labs(title = "Counts of Pet Species in Seattle", x = "Species", y = "Count") +
  theme_minimal()

species_plot

```
Write your narrative here We use ggplot2 to create a bar plot displaying the counts of pet species. We specify species as the x-axis using aes(x = species), then use geom_bar() to create the bars. We add a title and labels for the axes using labs(), and apply a minimal theme with theme_minimal() to improve the appearance of the plot. 




`9. (2 points)
```{r}
top_10_names <- seattlepets %>% 
filter(animal_name %in% c( "Lucy"  , "Charlie" , "Luna" , "Bella" , "Max"    , 
                           "Daisy" , "Molly"   , "Jack" , "Lily"  , "Stella" ))
top_10_names
```
`a.	What does the above code chunk do?



`b. Plot the counts of the pet names (animal_name) in top_10_names
```{r}

```


`10.  (2 points)

`The below code plots the proportion of dogs with a given name versus the proportion of cats with the same name. The 20 most common cat and dog names are displayed. The diagonal line on the plot is the x = y line; if a name appeared on this line, the name’s popularity would be exactly the same for dogs and cats.

```{r eval = TRUE, echo = FALSE, fig.fullwidth = TRUE}
# data prep
name_props <- seattlepets %>% 
  filter(
    !is.na(animal_name),
    species %in% c("Dog", "Cat")
    ) %>%
  group_by(species) %>% 
  count(animal_name, sort = TRUE) %>%
  mutate(prop = n / sum(n))
    
cat_name_props <- name_props %>%
  filter(species == "Cat") %>%
  rename(cat_prop = prop) %>%
  slice(1:30)

dog_name_props <- name_props %>%
  filter(species == "Dog") %>%
  rename(dog_prop = prop) %>%
  slice(1:30)
    
comb_name_props <- inner_join(cat_name_props, dog_name_props, 
                              by = "animal_name") %>%
  ungroup() %>%
  select(animal_name, cat_prop, dog_prop)
    
# create viz
ggplot(comb_name_props, aes(x = cat_prop, y = dog_prop)) +
  geom_abline(intercept = 0, 
              color = COL["lgray","full"], 
              alpha = 0.8, 
              size = 1.5) +
  geom_text_repel(aes(label = animal_name), 
                  segment.color = COL["gray","full"], 
                  seed = 291252, max.iter = 10000) +
  geom_point(color = COL["blue","full"], alpha = 0.8) +
  theme_minimal() +
  labs(x = "Proportion of cats", y = "Proportion of dogs") +
  xlim(0.002, 0.01) +
  ylim(0.002, 0.01) +
  ggimage::geom_emoji(
      image = "1f436", aes(x = 0.003, y = 0.009), size = 0.1
      ) +
  ggimage::geom_emoji(
      image = "1f431", aes(x = 0.009, y = 0.003), size = 0.1
      )
```

`What names are more common for cats than dogs? The ones above the line or the ones below the line?

`Answer here..................



`Is the relationship between the two variables (proportion of cats with a given name and proportion of dogs with a given name) positive or negative? What does this mean in context of the data?

`Answer here .............

